Capstone Movielens Report

### 2025-05-27 reports/templates/02.8.UMGYDE-model-template.Rmd
Started working on new model: model template added.

If we define $d_{i,j}$ as the _number of days since the earliest record_ in the `edx` dataset for movie $j$ rated by user $i$, then the formula \@res(eq:UMGYE-model) describing the _UMGYE Model_, for the current model, takes the form:

\begin{equation}
Y_{i,j} = \mu + \alpha_i + \beta_j + g_{i,j} + \gamma(v_{i,j}) + s(d_{i,j}) \varepsilon_{i,j}
(\#eq:UMGYDE-model)
\end{equation}

with $s$ a smooth function of $d_{i,j}$

Therefore, the formula \@res(eq:year-effect) for calculation the prediction of a _year effect_ as a residual, for a _smoothed day effect_ 
$$
\hat{f} = s(d_{i,j}) + \varepsilon_{i,j}
$$

@@

\begin{equation}
\hat{s(d_{i,j})} = Y_{i,j} - (\mu + \alpha_i + \beta_j + g_{i,j} + \gamma(v_{i,j}))
(\#eq:smoothed-day-effect)
\end{equation}

@@

@@

Below, we provide the most significant line of the code for training our model using the `5-Fold Cross Validation` method:
```{r, eval=FALSE}
```
```{r}
```

@@

```{r, eval=FALSE}
```
```{r}
RMSE_kable(RMSEs.ResultTibble.UMGYDE)
```

@@

We have already explained the idea of _Linear Model Regularization_ in the [UME Model Regularization] section above. We have also seen how the formula \@res(eq:ME-penalty) for adding a penalty to the _UME Model_ is transformed into the formula \@res(eq:GE-penalty) for the _UMGE Model_ and then into the formula \@res(eq:GYE-penalty) for the _UMGYE Model_. For the current model, this formula takes the form:

\begin{equation}
\sum_{i,j} \left(y_{u,i} - \mu - \alpha_i - \beta_j - g_{i,j} - \gamma(v_{i,j}) - s(d_{i,j})\right)^2 + \lambda \sum_{i,j} s(d_{i,j})^2
(\#eq:DE-penalty)
\end{equation}

And the formula \@res(eq:ME-regularized) for calculating the values of the _treatment effect_ that minimizes the equation will take the form:

\begin{equation}
\hat{s}(d_{i,j}, \lambda) = \frac{1}{\lambda + n_d} \sum_{r=1}^{n_{d}} \left(Y_{i,j} - \mu - \alpha_i - \beta_j - g_{i,j} - \gamma(v_{i,j})\right)
(\#eq:DE-regularized)
\end{equation}

@@

This function calculates _RMSE_ of the _UMGYDE Model_ using _5-Fold Cross Validation_ for the given $\lambda$ parameter value:
```{r, eval=FALSE}
```

@@

Below we provide the most significant part of the code that performs this operation:
```{r, eval=FALSE}
```
```{r}
str(cv.UMGYDE.preset.result)
cv.UMGYDE.preset.result$best_result
```

@@

Below we provide the most significant part of the code that performs this operation:
```{r, eval=FALSE}
```
```{r}
str(UMGYDE.rglr.fine_tune.results)
UMGYDE.rglr.fine_tune.results$best_result
```

@@

```{r, eval=FALSE}

@@

We calculate the _Root Mean Squared Error_ for the ultimately computed _UMGYDE Effect_ using the helper function [calc_UMGYD_effect_RMSE.cv](#func.calc_UMGYD_effect_RMSE.cv) described above in the [UMGYDE Model Regularization: Support Functions] section as follows:
```{r, eval=FALSE}
```

Finally, we add the definitive result for the current model to our _Result Table_:
```{r, eval=FALSE}
```
```{r}
```

>>>>

If we define $d_{i,j}$ as the _number of days since the earliest record_ in the `edx` dataset for movie $j$ rated by user $i$, then the formula \@ref(eq:UMGYE-model) describing the _UMGYE Model_, for the current model, takes the form:

\begin{equation}
Y_{i,j} = \mu + \alpha_i + \beta_j + g_{i,j} + \gamma(r_{i,j}) + f(d_{i,j}) \varepsilon_{i,j}
(\#eq:UMGYDE-model)
\end{equation}

with $f$ a smooth function of $d_{i,j}$

Therefore, the formula \@ref(eq:year-effect) for calculation the prediction of a _year effect_ as a residual, for a _smoothed day effect_ 
$$
\hat{f} = f(d_{i,j}) + \varepsilon_{i,j}
$$

@@

\begin{equation}
\hat{f(d_{i,j})} = Y_{i,j} - (\mu + \alpha_i + \beta_j + g_{i,j} + \gamma(r_{i,j}))
(\#eq:smoothed-day-effect)
\end{equation}

@@

#### [calc_date_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L2) Function {#func.calc_date_general_effect}
\

This is a helper function that calculates the _date general effect_ of the user ratings and is intended to be used in other functions and scripts (for instance in [train_UMGYD_effect](#func.train_UMGYD_effect)) described below.

The following is a slightly simplified version of the function's source code:
```{r eval=FALSE}
calc_date_general_effect <- function(train_set, lambda = 0){
  if (is.na(lambda)) {
    stop("Function: calc_date_general_effect
`lambda` is `NA`")
  }

  if(lambda == 0) put_log("Function `calc_date_general_effect`:
Computing Date Global Effect for given Train Set data...")
  else put_log1("Function `calc_date_general_effect`:
Computing Date Global Effect for lambda: %1...",
                lambda)
  dg_effect <- train_set |> 
    left_join(edx.user_effect, by = "userId") |>
    left_join(rglr.UM_effect, by = "movieId") |>
    left_join(rglr.UMG_effect, by = "movieId") |>
    left_join(date_days_map, by = "timestamp") |>
    mutate(resid = rating - (mu + a + b + g)) |>
  group_by(days) |>
    summarise(de = mean_reg(resid, lambda), 
              year = mean(year))
  dg_effect
}
```

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_date_general_effect](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L2) function is available in the [UMGYDE Model Support Functions](#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L2) script on _GitHub_.
:::

A data structure containg information about _Date General Effect_.

#### [calc_date_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L30) Function {#func.calc_date_general_effect.cv}
\

This is a helper function that calculates the _date general effect_ of the user ratings using the _5-Fold Cross Validation_ method and is intended to be used in other functions and scripts (for instance in [train_UMGYD_effect.cv](#func.train_UMGYD_effect.cv)) described below.

##### Souurce Code
\

::: {.noteblock data-latex=""}
The complete version of the source code of the [calc_date_general_effect.cv](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L30) function is available in the [UMGYDE Model Support Functions](#L1) section of the [UMGYD-effect.functions.R](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L30) script on _GitHub_.
:::

The following is a simplified version of the function's source code:
```{r eval=FALSE}
calc_date_general_effect.cv <- function(lambda = 0){
  if (is.na(lambda)) {
    stop("Function: calc_date_general_effect.cv
`lambda` is `NA`")
  }

  date_general_effect_ls <- lapply(edx_CV,  function(cv_fold_dat){
    cv_fold_dat$train_set |> calc_date_general_effect(lambda)
  })
  str(date_general_effect_ls)

  date_general_effect_united <- union_cv_results(date_general_effect_ls)

  date_general_effect_united |>
    group_by(days) |>
    summarise(de = mean(de, na.rm = TRUE), year = mean(year, na.rm = TRUE))
}
```

::: {.noteblock data-latex=""}
Here we use the function call [union_cv_results](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/data.helper.functions.R#L432) to aggregate the _5-Fold Cross Validation_ method results (for details, see the [union_cv_results](#func.union_cv_results) function description in the [Data Helper Functions](#appndx_a.data_helper.functions) section of the [Appendix] to this report).
:::

A data structure containg information about _Date General Effect_ as result of _5-Fold Cross Validation_.

@@

Below, we provide the most significant line of the code for training our model using the `5-Fold Cross Validation` method:
```{r, eval=FALSE}
  cv.UMGYD_effect <- train_UMGYD_effect.cv()
```
```{r}
str(cv.UMGYD_effect)
summary(cv.UMGYD_effect)
```

@@

```{r, eval=FALSE}
cv.UMGYD_effect.RMSE <- calc_UMGYD_effect_RMSE.cv(cv.UMGYD_effect)

RMSEs.ResultTibble.UMGYDE <- RMSEs.ResultTibble.rglr.UMGYE |> 
  RMSEs.AddRow("UMGYDE Model", 
               cv.UMGYD_effect.RMSE,
               comment = "User+Movie+Genre+Year+Day Effect (UMGYDE) Model")
```
```{r}
RMSE_kable(RMSEs.ResultTibble.UMGYDE)
```

@@

We have already explained the idea of _Linear Model Regularization_ in the [UME Model Regularization] section above. We have also seen how the formula \@ref(eq:ME-penalty) for adding a penalty to the _UME Model_ is transformed into the formula \@ref(eq:GE-penalty) for the _UMGE Model_ and then into the formula \@ref(eq:GYE-penalty) for the _UMGYE Model_. For the current model, this formula takes the form:

\begin{equation}
\sum_{i,j} \left(y_{u,i} - \mu - \alpha_i - \beta_j - g_{i,j} - \gamma(r_{i,j}) - f(d_{i,j})\right)^2 + \lambda \sum_{i,j} f(d_{i,j})^2
(\#eq:DE-penalty)
\end{equation}

And the formula \@ref(eq:ME-regularized) for calculating the values of the _treatment effect_ that minimizes the equation will take the form:

\begin{equation}
\hat{\gamma}(r, \lambda) = \frac{1}{\lambda + n_d} \sum_{i=1}^{n_{d}} \left(Y_{i,j} - \mu - \alpha_i - \beta_j - g_{i,j} - \gamma(r)\right)
(\#eq:DE-regularized)
\end{equation}

@@

This function calculates _RMSE_ of the _UMGYDE Model_ using _5-Fold Cross Validation_ for the given $\lambda$ parameter value:
```{r, eval=FALSE}
regularize.test_lambda.UMGYD_effect.cv <- function(lambda){
  if (is.na(lambda)) {
    stop("Function: regularize.test_lambda.UMGYD_effect.cv
`lambda` is `NA`")
  }

  train_UMGYD_effect.cv(lambda) |>
    calc_UMGYD_effect_RMSE.cv()
}
```

@@

Below we provide the most significant part of the code that performs this operation:
```{r, eval=FALSE}
  lambdas <- seq(0, 512, 32)
  cv.UMGYDE.preset.result <- 
    tune.model_param(lambdas, 
                     regularize.test_lambda.UMGYD_effect.cv,
                     steps.beyond_min = 16)
```
```{r}
str(cv.UMGYDE.preset.result)
cv.UMGYDE.preset.result$best_result
```

@@

Below we provide the most significant part of the code that performs this operation:
```{r, eval=FALSE}
endpoints <- 
  get_fine_tune.param.endpoints(cv.UMGYDE.preset.result$tuned.result)

UMGYDE.loop_starter <- c(endpoints["start"], 
                        endpoints["end"], 
                        8)
UMGYDE.loop_starter
#> [1]

cache.base_name <- "UMGYDE.rglr.fine-tuning"

UMGYDE.rglr.fine_tune.results <- 
  model.tune.param_range(UMGYDE.loop_starter,
                         UMGYDE.rglr.fine_tune.cache.path,
                         cache.base_name,
                         regularize.test_lambda.UMGYD_effect.cv)
```
```{r}
str(UMGYDE.rglr.fine_tune.results)
UMGYDE.rglr.fine_tune.results$best_result
```

@@

```{r, eval=FALSE}
  rglr.UMGYD_effect <- edx |> train_UMGYD_effect(UMGYDE.rglr.best_lambda)

@@

We calculate the _Root Mean Squared Error_ for the ultimately computed _UMGYDE Effect_ using the helper function [calc_UMGYD_effect_RMSE.cv](#func.calc_UMGYD_effect_RMSE.cv) described above in the [UMGYDE Model Regularization: Support Functions] section as follows:
```{r, eval=FALSE}
  rglr.UMGYD_effect.RMSE <- calc_UMGYD_effect_RMSE.cv(rglr.UMGYD_effect)
```

Finally, we add the definitive result for the current model to our _Result Table_:
```{r, eval=FALSE}
RMSEs.ResultTibble.rglr.UMGYDE <- RMSEs.ResultTibble.UMGYDE |> 
  RMSEs.AddRow("Regularized UMGYDE Model", 
               rglr.UMGYD_effect.RMSE,
               comment = "Computed for `lambda` = %1" |>
                 msg.glue(UMGYDE.rglr.best_lambda))

```
```{r}
RMSE_kable(RMSEs.ResultTibble.rglr.UMGYDE)
```
